# Results

```{r}
library(tidyverse)
library(ggplot2)
source("resources/plot_missing.R")
```

## Race 

```{r}
data <- read_csv("data/nyts_grouped.csv", show_col_types = FALSE)
```

```{r, eval=FALSE}
sum(data[5], na.rm = TRUE) # Not Hispanic or Latino
sum(data[6], na.rm = TRUE) # Mexican
sum(data[7], na.rm = TRUE) # Puerto Rico
sum(data[8], na.rm = TRUE) # Cuba
sum(data[9], na.rm = TRUE) # Another Hispanic or Latino
sum(data[10], na.rm = TRUE) # American Indian or Alaska Native
sum(data[11], na.rm = TRUE) # Asian
sum(data[12], na.rm = TRUE) # Black or African American
sum(data[13], na.rm = TRUE) # Native Hawaiian or Other Pacific Islander
sum(data[14], na.rm = TRUE) # White
```

```{r, eval=FALSE}
sum(data[5][data[15] == 1], na.rm = TRUE) # Not Hispanic or Latino
sum(data[6][data[15] == 1], na.rm = TRUE) # Mexican
sum(data[7][data[15] == 1], na.rm = TRUE) # Puerto Rico
sum(data[8][data[15] == 1], na.rm = TRUE) # Cuba
sum(data[9][data[15] == 1], na.rm = TRUE) # Another Hispanic or Latino
sum(data[10][data[15] == 1], na.rm = TRUE) # American Indian or Alaska Native
sum(data[11][data[15] == 1], na.rm = TRUE) # Asian
sum(data[12][data[15] == 1], na.rm = TRUE) # Black or African American
sum(data[13][data[15] == 1], na.rm = TRUE) # Native Hawaiian or Other Pacific Islander
sum(data[14][data[15] == 1], na.rm = TRUE) # White
```

```{r}
race_ethnicity = c('Mexican', 'Puerto Rico', 'Cuba', 'Another Hispanic or Latino', 'American Indian or Alaska Native', 'Asian', 'Black or African American', 'Native Hawaiian or Other Pacific Islander', 'White')

total_race_count = c(13827, 2895, 1081, 10074, 7182, 6684, 18918, 2912, 56445)

smoked_per_race = c(2522, 508, 267, 1523, 1502, 747, 2958, 503, 9882)

race_df = data.frame(race_ethnicity, total_race_count, smoked_per_race)
```

```{r}
race_df <- race_df %>%
  mutate(percent_smoking = smoked_per_race / total_race_count) %>%
  arrange(percent_smoking)

race_bar <- race_df %>%
  #arrange(percent_smoking) %>% 
  ggplot(aes(x=reorder(race_ethnicity, desc(percent_smoking)), y=percent_smoking)) +
  geom_bar(color='blue', fill = 'lightblue', stat='identity') +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  xlab('Race/Ethnicity') +
  ylab('% Tried Smoking') +
  ggtitle('Percentage of Youth That Has Tried Smoking by Race/Ethnicity')

race_bar
```


As we can see by the breakdown by race/ethnicity above, almost 1 in every 4 of the youth that identify as Cuban have tried smoking. The percentage of youth identifying as Cuban who have tried smoking is over twice as much as the percentage of youth who identify as Asian. The percentage of the youth who identify as Hispanic or Latino and have tried smoking is about 17%, which is greater than most other race/ethnicities other than American Indian or Alaska Native. Approxmately 1 in 5 of the youth who identity as American Indian or Alaska Native have tried smoking. 



